---
title: "dymiumCore for modellers"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{dymiumCore for modellers}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, echo = FALSE, message=FALSE}
library(dymiumCore)
```

# Introduction

This manual is for people who would like to put together a microsimulation model using ready-to-use modules from [dymium-org/dymiumModules](https://github.com/dymium-org/dymiumModules). The most the you are willing to change are entity data and model parameters that go into your microsimulation model. If this is you then read on! :) otherwise please see the manual section and select the appropriate manual for you.

# Requirements

It is recommended that you setup a new RStudio project for your microsimulation project. See this [page](https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects) if you are using RStudio but new to using Projects. If you don't have RStudio please go ahead and download it at (rstudio.com).

Every module in [dymium-org/dymiumModules](https://github.com/dymium-org/dymiumModules) repository rely on the [modules](https://cran.r-project.org/web/packages/modules/index.html) package to work properly. Hence, make sure your have the package installed.

# Download and use existing modules

I recommend that before you download any module you should first see its README page inside the repository. For example, if you want to know about the `demography` module or the events it has you should visit https://github.com/dymium-org/dymiumModules/tree/master/modules/demography.

Once you are certain which module you want then use `download_module()`. This function allows modules to be downloaded into your active RStudio project. For example, if you would like to download the `matsim` module:

```{r, eval=FALSE}
download_module("demography")
```

By default, this downloads the latest version of the `demography` module from [dymium-org/dymiumModules](https://github.com/dymium-org/dymiumModules) into a newly created folder called `modules` at the root of your active RStudio project. See `?download_module` for more download options.

# Use an event

Inside the __demography__ module you will find a number of R scripts. `helpers.R`, `constants.R`, `logger.R`, `tests` folder are files and folder that got created when a new module is created using `use_module()`. To learn more about them see the developer manual. The R scripts other than those are event scripts and they are the ones that you should import to your microsimulation model.

For example, if you would like to use the birth event which can be found at `your_project_folder/modules/demography/birth.R` in your microsimulation model do as the following:

```{r, eval = FALSE}
event_demography_birth <- modules::use('modules/demography/birth.R')
```

The above chuck imports the birth event into your active R environment by assigning it to a variable called `event_demography_birth`. Any event that is imported using `modules` have the following fields exposed: `run()` and `REQUIRED_MODELS`. 

```{r, eval = FALSE}
event_demography_birth$run()
```

The `run()` function of any event takes the same four main arguments which are `x`, `model`, `target`, `time_step`. Some `run()` functions may take more than the four main arguments. Therefore you should always check the README page of the module of the event that you are using.

```{r, eval = FALSE}
event_demography_birth$REQUIRED_MODELS
```

`REQUIRED_MODELS` is a field that contains `NULL` if no models are required by the event or a character vector. Anytime `run()` is called it will check if the supplied model argument or the supplied world has all the required models or not. If not then an error will appear.


# Let's put together a microsimulation model


```{r}

```


