---
title: "Mate matching"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Mate matching}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  out.width = '100%',
  fig.width=7, 
  fig.height=5, 
  dpi=150
)
```

```{r setup}
# To install a package use `install.packages('package-name')`
library(dymiumCore)
library(data.table)
library(R6)
library(dplyr)
library(ggplot2)
library(gganimate)
set.seed(728)
knitr::opts_chunk$set(echo = TRUE)
```

## Load population

```{r, asis = TRUE}
# Import individual and household data
ind_data <- toy_individuals

# Create an Individual object with attribute data
Ind <- Individual$new(.data = ind_data, id_col = "pid")

# show data
Ind$get_data()
```
## Randomly assign group to agents

All agents will be randomly assigned a group number to them this is to create 
sub markets. Sub markets are defined such that only agents within the same sub 
market can interact with each other. This maybe used to mimic agents' social 
network, geographical boundary, etc. 

```{r}
n_groups <- 2

# randomly assign group numbers to all agents
Ind$get_data(copy = FALSE)[, group := sample(paste0("Group_", 1:n_groups), size = .N, replace = TRUE)]
```

```{r}
Ind$get_data() %>%
  ggplot(data = ., aes(x = age, fill = sex)) +
    geom_density() +
    facet_grid(group ~ sex) +
    ggthemes::scale_fill_few() +
    ggthemes::theme_few(base_size = 12) +
    labs(title = "Distribution of age by gender and group.")
```


## Implement matching markets 

New matching market can be extended from the available matching market classes 
which are `MatchingMarketStochastic` and `MatchingMarketOptimal`. In every matching
problems there are two sides of the market, we call them A and B. Scoring functions
need to be defined for each side of the market manually by implementing `matching_score_A`
and `matching_score_B`. `matching_score_A` represent how agents in one side of
the market evaluate the candidates of the opposite side, and vice versa for 
`matching_score_B`. The user is expected to change the function but the three 
main argument for both the scoring functions must remain the same. `self$matching_problem`
allows the function to access the variables given when the class was created. 

The rule that all agents follow is simple, they want to find a match that has 
the least age difference, as expressed in the scoring functions `matching_score_A` 
and `matching_score_B`.

$$ MatchingScore_x = \frac{1}{1+|age_x - age_y|}$$
```{r}
require(lattice)
x <- seq(-10, 10, length.out = 15)  
y <- x  
f <- function(x,y) {
  1 / (1 + abs(x-y))
}
z <- outer(x, y, f)  
wireframe(z, drape=T, col.regions=rainbow(100), zlab = "Score", xlab = "age_x", y = "age_y")
```


```{r}
MateMatchingStochastic <- R6::R6Class(
  classname = "StochasticCohabitationMarket",
  inherit = MatchingMarketStochastic,
  public = list(
    matching_score_A = function(matching_problem = self$matching_problem,
                                idx_A,
                                idx_B) {
      scores <- 1 / (1 + abs(
        matching_problem$agentset_A[["age"]][idx_A] -
          matching_problem$agentset_B[["age"]][idx_B]
      ))
    },
    matching_score_B = function(matching_problem = self$matching_problem,
                                idx_B,
                                idx_A) {
      scores <- 1 / (1 + abs(
        matching_problem$agentset_B[["age"]][idx_B] -
          matching_problem$agentset_A[["age"]][idx_A]
      ))
    }
  )
)
```


```{r}
MateMatchingOptimal <- R6::R6Class(
  classname = "MateMatchingOptimal",
  inherit = MatchingMarketOptimal,
  public = list(
    matching_score_A = function(matching_problem = self$matching_problem, idx_A, idx_B) {
      outer(X = matching_problem$agentset_B[["age"]][idx_B], # reviewers, rows
            Y = matching_problem$agentset_A[["age"]][idx_A], # proposers, columns
            function(x, y) {
              1 / (1 + abs(x - y))
            })
    },
    matching_score_B = function(matching_problem = self$matching_problem, idx_B, idx_A) {
      outer(X = matching_problem$agentset_A[["age"]][idx_A], # reviewers, rows
            Y = matching_problem$agentset_B[["age"]][idx_B], # proposers, columns
            function(x, y) {
              1 / (1 + abs(x - y))
            })
    }
  )
)
```

## Simulation

```{r}
# using dplyr's filter
agentset_A <- Ind$get_data() %>%
  dplyr::filter(age > 20 & age < 60 & sex == 'male') %>%
  data.table::setDT(.) # convert back to data.table
  
# using data.table syntax for filtering
agentset_B <- Ind$get_data() %>%
  .[age %between% c(20, 60) & sex == 'female']

# define id column
id_col_A <- id_col_B <- 'pid'

# create a market object using its constructor
StochasticMatching <- MateMatchingStochastic$new(agentset_A, agentset_B, id_col_A, id_col_B, grouping_vars = c('group'))

OptimalMatching <- MateMatchingOptimal$new(agentset_A, agentset_B, id_col_A, id_col_B, grouping_vars = c('group'))
```

```{r}
stochastic_matching_result <- 
  StochasticMatching$simulate(n_choices = 10, # number of potential partners each agent evaluates
                              method = 'pweighted', # weighted probability
                              by_group = TRUE) # segments the whole population into by markets by their group

optimal_matching_result <- 
  OptimalMatching$simulate(method = "one-to-one", # each agent only has one match.
                           one_sided = FALSE, # agents from both sides of the market have preferences towards each other
                           optimal_A = TRUE, # optimal for A
                           by_group = TRUE) # segments the whole population into by markets by their group
```

## Visualise the match qualities of the two approaches.

```{r}
stochastic_matching_quality <- 
  na.omit(stochastic_matching_result) %>%
  merge(x = ., y = Ind$get_data()[, .(pid, age)], by.x = 'id_A', by.y = 'pid') %>%
  merge(x = ., y = Ind$get_data()[, .(pid, age)], by.x = 'id_B', by.y = 'pid', suffixes = c('_A', '_B')) %>%
  .[, age_diff := age_A - age_B]

ggplot(data = stochastic_matching_quality, aes(age_diff)) +
  geom_density(color = "black", fill = "steelblue") +
  theme_bw() +
  labs(title = "A density plot of age differences in the stochastic mate matching market.")

optimal_matching_quality <- 
  na.omit(optimal_matching_result) %>%
  merge(x = ., y = Ind$get_data()[, .(pid, age)], by.x = 'id_A', by.y = 'pid') %>%
  merge(x = ., y = Ind$get_data()[, .(pid, age)], by.x = 'id_B', by.y = 'pid', suffixes = c('_A', '_B')) %>%
  .[, age_diff := age_A - age_B]

ggplot(data = optimal_matching_quality, aes(age_diff)) +
  geom_density(color = "black", fill = "salmon") +
  theme_bw() +
  labs(title = "A density plot of age differences in the optimal mate matching market.")
```

