---
output:
  github_document:
    toc: true
    toc_depth: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)
```

# {{{module}}} module documentation

- Authors: Your name
- Email: your@email.com
- Current version: 0.1

This module provides the following functionalities...

# Release notes

## Version 0.1

This is the first version of this module.

# Requirements

This module requires the following packages ... and external dependecies such as ... 

# Events

## Event name 1

### Usage

```{r}
event_{{{module}}}_eventname <- modules::use("{{{module_path}}}/eventname.R")
event_{{{module}}}_eventname$run(world, model = NULL, target = NULL, time_steps = NULL)
```

### Params

* world: a World object.
* model: a named list.

```{r}
# For example
model <- list(x1 = list(yes = 0.5, no = 0.5),
              x2 = data.table(sex = c("male", "female"), prob = c(0.2, 0.1)))
```

* target: a named list.
* time_steps: a numeric vector that indicates the time steps in the simulation clock
that this event should be run.

### Description

This event does..

### Note

Here are some assumptions and settings that you should know about this event..

### Example

```{r}
1 + 1
```

## Event name 2

### Usage

```{r}
event_{{{module}}}_eventname <- modules::use("{{{module_path}}}/eventname.R")
event_{{{module}}}_eventname$run(world, model = NULL, target = NULL, time_steps = NULL)
```

### Params

* world: a World object.
* model: a named list.

```{r}
# For example
model <- list(x1 = list(yes = 0.5, no = 0.5),
              x2 = data.table(sex = c("male", "female"), prob = c(0.2, 0.1)))
```

* target: a named list.
* time_steps: a numeric vector that indicates the time steps in the simulation clock
that this event should be run.

### Description

This event does..

### Note

Here are some assumptions and settings that you should know about this event..

### Example

```{r}
1 + 1
```

# Known issues

Here are some known issues
